### 2021-03-14 paper/rebuttal/rebuttal.tex

\rcomment{%
	Another interesting point in this work is that the authors chose to not report p-values or the results of statistical significance tests (except confidence intervals). I presume this was a conscious decision, which I do not necessarily agree with. I see no critical error in the approach used by the authors, but I would have found aspects of the manuscript easier to understand in places with results of statistical tests (e.g., associated with goodness-of-fit tests formally test that the embedding-based distances are more correlated with the flux values). Given the sample sizes involved, I have no doubt that the results are statistically significant. However, I personally do see value in rigorously performed frequentist hypothesis testing, and I felt that this manuscript could be improved by employing such techniques.

	Lastly, the authors do not actually state how the 99 confidence intervals were constructed. (E.g., was this via a bootstrap approach?) This detail should be added.
}

\response{Thank you for this feedback.
	You are correct that it was a conscious choice to not report statistical significance in the form of a $p$-value.
	The use of statistical significance is contentious, and Nature itself has hosted some of this debate~\autocite{amrhein2019statistics}.
	As such, we avoided $p$-values in our research when it made sense to do so.
	In our analysis, all significance levels were below 0.0001, with most computationally indistinguishable from zero.
	However, we recognize that $p$-values can help readers interpret results.
	As such, we have opted to add to a sentence such as \say{\textit{Correlation is calculated on the data in the log-log scale ($p < 0.0001$ across all fits)}.} or similar to relevant figures in the main text and supporting materials.

	We have also provided information on how confidence intervals were calculated in relevant figure captions.
	For example, we state in figure 2's caption that \say{\textit{99\% confidence intervals are plotted for the mean flux in each bin based on a normal distribution.}}, which should give readers enough information to reproduce the confidence intervals following their standard definition.
	Other confidence intervals, such as intervals for the loess regression line in figure 6, are calculated using the smoothing functions in the \texttt{ggplot2} package of R and we refer readers to the relevant documentation.
}

@@

\rcomment{%
	F. Suggested improvements:
	References and discussion of prior work using word2vec-style embeddings of mobility data.
	Additional experiments using alternative embedding methods, especially an SVD-based embedding method, as well as robustness/reproducibility studies across different runs of the SGNS would be essential to strengthen the paper.
	In addition, some additional results demonstrating how the embeddings provide novel scientific insight, rather than just recapitulating known structures, would be essential to make this work above-the-bar for a high-impact publication such as Nature.

}

\response{We believe that our new mathematical results substantially improve the manuscript, extends its contribution, demonstrates novel scientific insight, and justifies the use of \textit{word2vec} towards mobility data. By adding this, and addressing your other points (above), we believe that the revised manuscript sufficiently remedies your concerns.

@@

\rcomment{%
	Although the paper is well written and explores a very interesting problem, I have certain criticism for the used methodology. After obtaining a representation of each dataset, the authors defined an embedding distance as the cosine distance between vectors in the space, while defining distance as the geographical distance in the original, spatially embedded datasets. Subsequently they used the Gravity Law to estimate the flux between location pairs of vector pairs in the spatial and embedded space respectively. The Gravity Law, applied on mobility, assumes that the number of people moving between two populations is proportional to the multiplicative of the sizes of the populations and a deterrence function. Research during the last decades on transportation, spatial networks and human mobility suggests that this deterrence function may take a functional form of an inverse exponential or an inverse power-law, as it has been acknowledged by the authors. However, the emergence of these different deterrence functions can be due to fundamentally different mechanisms (for a review on the topic see Barthélemy, Physics Reports 499, 1, 2011, a paper not cited in the manuscript). For example, the different cost functions (linear or logarithmic) to create link at a given distance, the spatial granularity of the observation, or even the heterogeneity of the moving individuals in terms of various factors, just to mention a few examples. Instead, the authors of the actual paper arbitrarily assumed a power-law deterrence function in the spatial case and an exponential deterrence function in the embedding space without further arguing their choice (other than fitting on the data). Furthermore, assuming different deterrence functions for the flux estimation using Gravity Law makes the estimated fluxes incomparable, resulting the invalidation of results depicted in Figure 2.
}

\response{Thank you for noting this.
	We carefully read out the Barthélemy, Physics Reports 499, 1, 2011, and found it incredibly educational, referencing it in our manuscript.		
	Past work has argued that the power-law decay function is appropriate for modelling geographic distance---being more suitable for the analysis of large, complex, and scale-free spatial systems with long-range spatial memory---whereas the exponential gravity model fits well with the simpler system with no spatial memory~\autocite{chen2015distance}, such as might be encoded in the \textit{word2vec} embedding space.
	We report the performance of the flux estimation and prediction using both deterrence functions (see Section 2, above), finding that although the absolute values vary, the qualitative pattern remains the same---the embedding distance outperforms geographic distance regardless of the function used.
	 We also elaborate on the connection between the embedding distance and exponential model in our discussion on the relationship between \textit{word2vec} and the gravity model of mobility (see Section 1, above).
	 In addition, we add the following line to our manuscript, noting these considerations,

	\medskip
	\say{The decay function $f(r_{ij})$ defines the effect of distance, and so different decay functions can model fundamentally different mechanisms [...] such as the cost functions for a given distance and the spatial granularity of the observation.
		For geographic distance, we define $f(r_{ij})$ as the standard power-law function, and for the embedding distance, we use the exponential function, selected as the best performing for each case though there is no qualitative difference---the embedding distance outperforms geographic distance regardless of the type of gravity model used (see Supporting Information).}

@@

}
\response{
	Thank you for these comments.
	First, we added a line to the methods section detailing the \textit{word2vec} we used, and noting that these achieved the best performance: \say{\textit{For main results regarding scientific mobility, we used $d=300$ and $w=1$, which were the parameters that best explained the flux between locations, though results were robust across different settings (Fig....)}}.

	A full search of the hyper-parameter space would certainly be interesting and useful, however we believe that set of hyper-parameters we considered here is sufficient to establish the utility of the approach, and demonstrate that performance is only slightly different based on different hyper-parameters.
	There are previous studies on the optimal dimensionality of word embedding \autocite{yin2018understand, yin2018dimensionality}. Also,
	we note that in previous work from our group, we explored the optimal dimensions for graph embedding~\autocite{gu2020dimensions}, finding that while a sufficient number of dimensions is necessary to capture the structure, increasing the dimensions does not heavily deteriorate the performance, a property common across deep learning methods.
	Given that there is little performance difference between $d = 100$ and $d = 300$, we opt to use the larger embeddings at $d = 300$, which are large enough to capture the complexity of mobility, yet small enough to allow an efficient computational workflow.
}

\rcomment{
	When datasets in the intro are explained, it should be mentioned that the detailed descriptions are in the Methods section.
}
\response{Thank you. We have added a parenthetical to the introduction after stating the datasets, as follows:  \say{We embed trajectories from three massive datasets: U.S. passenger flight itinerary records, Korean accommodation reservations, and a dataset of scientists' career mobility between organizations captured in bibliometric records (Detailed descriptions are available in the Methods).}
}
\rcomment{
	Some minors comments
	\begin{enumerate}
		\item There is no reference to Fig.5e and f in the text.
		\item Second paragraph on page 10: the authors discuss Fig.6 here but missed to refer the figure.
		\item Page 15, last row: double point in the end of sentence.
		\item SI Page 2, S3: “. if” $\rightarrow$ “. If”
	\end{enumerate}
}
\response{Thank you, we have addressed these comments and have updated the manuscript accordingly,
}

\rcomment{
	SI Fig.S1c: Here the authors conclude that the proportions “decreased over time”. These proportions seem to be rather flat without strong decreasing trends. In any case, if the authors would like to indicate minor changes here, it would be helpful for the reader if a grid would be posted on the figure.
}
\response{Thank you, we agree that the trend is flat, rather than decreasing. We have adjusted the text accordingly, and also added a grid to the figure so that it can be read more easily.

@@

\response{Thank you for this insightful comment!
	We think that there is a great deal of potential in methods like BERT for representing mobility data.
	What initially drew us to \textit{word2vec}, however, was its simplicity, ubiquity, and mathematical tractability (at least compared to other, more sophisticated approaches).
	This last attribute proved very useful, as in exploring the mathematical properties of the model, we discovered a deep connection between \textit{word2vec} and the gravity model of mobility, which explains why it learns gravity-like relationships between entities (see Section 1, this letter).
	We believe that this new discovery, now included in the manuscript, strongly grounds our use of \textit{word2vec} here.

	We agree that the new revolutions in neural embeddings offer huge potential for the study of human mobility, however we argue that our contribution is dependent on \textit{word2vec}, both for its equivelance with the gravity model of mobility, as well as for the model's simplicity and ubiquity.
	\textit{Word2vec} is also efficient; deep neural models often require tremendous amounts of training data ($\sim$3 billion words in BERT) and model parameters, making them infeasible for most studies of mobility.
	We hope that this manuscript serves as a foundation that others can build upon, applying more and more sophisticated neural embedding techniques to gain deeper insights into mobility data.

	To highlight the importance of alternative models, however, we added the following line to our conclusion: \say{\textit{Future studies may also consider bi-directional embeddings, such as BERT~\autocite{devlin2018bert}, to incorporate directionality.}}.

@@

}

\rcomment{
	The paper does not compare human transition (or co-affiliation) data with any other kinds of
	data that could be used for embedding. Does author byline data do better than the embedding
	of collaborating institutions (almost certainly), than co-mention of the institutions in the press, as
	with institutional collaborations (e.g, the Harvard-MIT Broad Institute, or joint university
	conferences; these will be stronger signals, but there are fewer of them), than advisor-student
	university pairs (although many of these will be captured by the authors’ approach when
	students publish at their home institutions before securing jobs elsewhere).
}

\response{
	Thank you for this comment!
	We feel that applying these embeddings to other areas of scientific activity would be incredibly enlightening.
	Your recommendation of collaboration (based on the author byline) and advisor-student pairs are especially interesting, and we intend to approach such questions in future studies. 
	However, for this manuscript, we believe that mobility is the most appropriate area of study, due to the importance of mobility to science, economics, and innovation, as well as the fundamental mathematical relationship between \textit{word2vec} and the gravity model of mobility, which we discovered since our first submission. 
}

@@

\rcomment{
	There are known concerns with clustering a 300 dimensional objects because of the curse of
	dimensionality. Many have shown that reducing dimensionality with UMAP to a nontrivial
	number of dimensions (e.g., 30), which preserves both local and some global distance, then
	clustering on those dimensions are much more stable (e.g., this post). This was one
	concern with the stability of the clustering analysis

}

\response{Thank you for this suggestion.
	We looked into this, using UMAP to reduce the country-level embeddings from 300 down to 30 dimensions, however, the results were incomprehensible, and much inferior to those using 300 dimensions.
	Many dimensionality reduction techniques reduce the number of dimensions by maintaining pairwise distances between vectors, so that the result can be well-suited for clustering algorithms.
	However, our approach already uses raw pairwise similarities to perform hierarchical clustering, so approximating this distance in 30 dimensions, and then using the new distances to perform clustering is redundant.
	We also find that the curse of dimensionality is not a major issue in our analysis, at least given the size of our data and the number of locations represented---\textit{word2vec} seems able to learn dense vector-space representations.
}

@@

\rcomment{
	P6Para3: “For THEIR purpose…”
}

\response{Thank you, we have corrected this mistake.
}

\rcomment{
	P8Para1: “by explicitly the relative importance of different level of the hierarchy with a scaling parameter r ” which is never described or explained. Considering it was derived by the authors in a very recent paper, no one will know what it is.
}

\response{Thank you, we have edited the text to read as “We quantify the relative importance of geography (by region), and language (by the most widely-spoken language of each country) using the element-centric clustering similarity, a method that can compare hierarchical clustering and disjoint clustering (geography, language...) at the different levels of hierarchy by explicitly adjusting a scaling parameter $r$, acting like a zooming lens”
}

@@

\rcomment{
	My second group of questions pertain to the prestige hierarchy, that really drives the dynamics of scientific mobility. The technical description addressed most of them. However, I was wondering, is it all possible to go beyond the descriptive part, and show cases where the more accurate embedding base distances do make a difference? That is, using them, we do not only obtain better correlations with the ranking data, but also allows us to address problems that we could not without the embedding distance? For example, some results have shown that a paper coauthored by two authors at different high prestige institutions has higher impact, than writing a paper with two authors of the same high prestige institution—would embedding help us pull apart such effects? Here I am not trying to make the authors to write a different paper, so maybe a little forward thinking may be enough, if doing the actual work is too time consuming, and goes beyond the current narrative.
}

\response{Thank you for this idea!
	We believe that this is an interesting topic and we hope to investigate it in the future.
	In your particular example, an embedding distance trained using sequences of co-authorship on the same paper might better capture the \textit{collaborative} distance between institutions.
	Then, using that distance, one could measure the ``embedding distance'' spanned by all authors in the byline and answer the question of whether typical collaborations relate to impact or not.
	This is certainly a topic that  we intend to pursue in a future study.
	In the case of scientific mobility, we might also consider the ``embedding distance'' spanned by an individual's career, and relate this to their demographics and scientific impact.

	However, while these topics are exciting, we believe that they fall outside the scope of the current manuscript, especially given the discoveries we made concerning a fundamental connection between the gravity law of mobility and \textit{word2vec}, which adds a substantial amount of new results to the manuscript.
}

@@

\rcomment{
	Finally, one cannot escape the fact that COVID changes our mobility patterns. Many groups have used mobile phone data to explore these changes. I wonder, how would the pre-covid embeddings apply to post-covid mobility? Once again, I am mindful that this would require access to such data, but I would not be surprised at all if the authors have access to such data—in which case this is remains an intriguing question.
}

\response{This is an excellent idea and one that we have been interested in addressing as well.
	Unfortunately, however, we do not have access to such data, at least not in large enough quantities to be representative.
	Publications are on a time lag, compared to actual events.
	For example, a researcher may write a paper while studying at university $A$, but then before the final paper is published, move to university $B$, so while they now have a $B$ affiliation, in our data all we see is the $A$ one.
	As such, we would not necessarily expect to see the impact of COVID on bibliometric data until later in 2021 or early 2022, when researchers have had the chance to publish with their new institution.
	We also note that some activity will, of course, be visible, especially given the number of researchers rapidly publishing on the subject of COVID-19, however, we consider this to be an unrepresentative sample, given that it is concentrated in a few disciplines.

@@

%
% R2 Table
%
\clearpage

\begin{landscape}

	\providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
	\providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
	\providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
	\providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}

	\begin{table}[ht]
		\caption{
			\textbf{Correlation between flux and distance over metrics, experimental parameters}.
			Each cell corresponds to the correlation between the real-world flux between scientific organizations (measured with $R^{2}$) and baseline metrics, shown by subsets of mobility data and by definitions of organization population.
			The asterisk denotes the top-performing distance metric by column.
			Distance metrics are ordered from highest $R^{2}$ to lowest, based on global mobility with organization population defined using all mobile and non-mobile authors.
			``All'' means that population is defined as the average yearly number of unique mobile and non-mobile scholars who published with the organizations' affiliation;
			population is defined in the same way for ``Mobile only'', except only using unique mobile researchers;
			``Raw freq'' means that organization populations are defined as their frequency across all the trajectories, similar to word frequency in language embedding.
			Embedding distance, measured as the cosine distance between embedding vectors, explains more of the flux than baselines in nearly every case, except using raw frequency population and domestic and international mobility, where direct optimization of the gravity model works better, as well as Levy's factorization~\autocite{levy2014neural} for domestic and international only mobility.
		}
		\label{supp:table:r2_table}
		\begin{centerbox}
			\begin{threeparttable}
				\small
				\setlength{\tabcolsep}{0pt}
				\begin{tabular}{l l l l l l l l l l}

@@

\end{landscape}

%
% RMSE TABLE --- Exponential model
%
\clearpage
\begin{landscape}

	\providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
	\providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
	\providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
	\providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}

	\begin{table}[ht]
		\caption{
			\textbf{Prediction error between actual and predicted mobility with exponential gravity model, by metrics and experimental parameters}.
			Each cell corresponds to the prediction error (measured with root mean squared error) when using each distance as input to the exponential form of the gravity model of mobility to predict the flux between organizations, shown by subsets of mobility data, and by definitions of organization population.
			The asterisk denotes the top-performing distance metric by column (lowest prediction error).
			Distance metrics are ordered from lowest prediction error to highest, based on global mobility with organization population defined using all mobile and non-mobile authors.
			``All'' means that population is defined as the average yearly number of unique mobile and non-mobile scholars who published with the organizations' affiliation;
			population is defined in the same way for ``Mobile only'', except only using unique mobile researchers;
			``Raw freq'' means that organization populations are defined as their frequency across all the trajectories, similar to word frequency in language embedding.
			Embedding distance, measured as the cosine distance between embedding vectors, results in better predictions of mobility than baselines in nearly every case, however Levy's factorization~\autocite{levy2014neural} perform better in the case of international and domestic only mobility when using raw frequency populations.
		}
		\label{supp:table:rmse_exp_table}
		\begin{centerbox}
			\begin{threeparttable}
				\small
				\setlength{\tabcolsep}{0pt}
				\begin{tabular}{l l l l l l l l l l}

@@

\end{landscape} % End r2 table page

%
% RMSE TABLE --- Power Law
%
\clearpage
\begin{landscape}

	\providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
	\providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
	\providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
	\providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}

### 2021-03-14 paper/rebuttal/rebuttal.tex

	\begin{table}[ht]
		\caption{
			\textbf{Prediction error between actual and predicted mobility with power-law gravity model, by metrics and experimental parameters}.
			Each cell corresponds to the prediction error (measured with root mean squared error) when using each distance as input to the power-law form of the gravity model of mobility to predict the flux between organizations, shown by subsets of mobility data, and by definitions of organization population.
			The asterisk denotes the top-performing distance metric by column (lowest prediction error).
			Distance metrics are ordered from lowest prediction error to highest, based on global mobility with organization population defined using all mobile and non-mobile authors.
			``All'' means that population is defined as the average yearly number of unique mobile and non-mobile scholars who published with the organizations' affiliation;
			population is defined in the same way for ``Mobile only'', except only using unique mobile researchers;
			``Raw freq'' means that organization populations are defined as their frequency across all the trajectories, similar to word frequency in language embedding.
			Embedding distance, measured as the cosine distance between embedding vectors, results in better predictions for global mobility with ``All'; and ``Mobile only'' definitions of population, though direct optimization of the gravity model with MDS performs better when using raw frequencies to measure organization population, and Levy's factorization~\autocite{levy2014neural} perform better here with domestic and international only mobility.
		}
		\label{supp:table:rmse_power_table}
		\begin{centerbox}
			\begin{threeparttable}
				\small
				\setlength{\tabcolsep}{0pt}
				\begin{tabular}{l l l l l l l l l l}
